### 2018-05-09 2_wmm.tex
Add some text, picture on compiler arch. cleanup a bit.

For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\mathr{fr} = (\mathr{rf}^{-1};\mathr{co})$. Figure~\ref{example:x86-cat} contains an example of x86-TSO weak memory model definition in \cat{} language that asserts acyclicity of communication relation, \rel{po-loc} relation, \rel{mfence} relation and some other derived relations~\cite{owens2009better}:

\begin{figure}[h]
\begin{lstlisting}
"X86 TSO"
include "x86fences.cat"
include "filters.cat"
include "cos.cat"

(* Uniproc check *)
let com = rf | fr | co
acyclic po-loc | com

(* Atomic *)
empty rmw & (fre;coe)

(* Global happens-before *)
#ppo
let po_ghb = WW(po) | RM(po)

#mfence
include "x86fences.cat"

#implied barriers
let poWR = WR(po)
let i1 = MA(poWR)
let i2 = AM(poWR)
let implied = i1 | i2

let com = rfe | fr | co
let ghb = mfence | implied | po_ghb | com
show implied
acyclic ghb as tso
\end{lstlisting}
\caption{The x86-TSO memory model defined in \cat{} language\protect\footnotemark}
\label{example:x86-cat}
\end{figure}
\footnotetext{The \cat{}-memory model for x86-TSO can be found in official repository of \tool{herd} tool: \url{https://github.com/herd/herdtools7}}
%\begin{figure}[h]
%\begin{lstlisting}
%...
%let com = rf | fr | co
%let po-loc = po & loc
%acyclic po-loc | com
%\end{lstlisting}
%\caption{Excerpt from the x86-TSO memory model in \cat{} language}
%\label{example:x86-cat}
%\end{figure}

%Relationship between different models \url{http://wiki.expertiza.ncsu.edu/index.php/CSC/ECE_506_Spring_2013/10c_ks}

%The work~\cite[Chapter~2]{alglave2010shared} provides formal definition of basic properties of relations, such as \textit{reflexivity} and \textit{irreflexivity}, \textit{transitivity} and \textit{transitive closure}, \textit{acyclicity}, that can be asserted by a weak memory model in order to restrict the set of allowed behaviours of the system.

>>>>

For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\mathr{fr} = (\mathr{rf}^{-1};\mathr{co})$. As an example memory model definition in \cat{} language, the the x86-TSO can be found in Appendix~\ref{apx:x86cat}. This memory model asserts acyclicity of communication relation, \rel{po-loc} relation, \rel{mfence} relation and some other derived relations~\cite{owens2009better}.

### 2018-05-09 4_impl.tex
Add some text, picture on compiler arch. cleanup a bit.

\begin{figure}[!h]%[!hb]
\centering
\begin{lstlisting}[language=antlr,
                  basicstyle=\ttfamily\tiny,
                  caption={The sketch of the input language grammar used by \oldporthos},%
                  label={fig:in_grammar_pts}]
<prog> : <init> <thrd>* <assert>
       ;
<thrd> : thread <tid> <inst>
       ;
<inst> : <atom>
       | <inst> ; <inst>
       | while <pred> <inst>
       | if <pred> { <inst> } <inst>
       ;
<atom> : <reg> <- <expr>
       | <reg> <:- <loc>
       | <loc> := <reg>
       | <reg> = <loc>.load(<atomic>)
       | <loc> = <reg>.store(<atomic>)
       | ('mfence' | sync' | 'lwsync' | 'isync')
       ;
<pred> :
       | true
       | false
       | <expr> (and | or) <expr>
       | <expr> ('==' | '!=' | '>' | '>=' | '<' | '>=') <expr>
       ;
<expr> : [0-9]
       | <reg>
       | <expr> ('*' | '+' | '-' | '/' | '%') <expr>
       ;
\end{lstlisting}
\end{figure}

@@

@@

The input memory model language \cat is discussed in Section~\ref{ch:wmm:cat}.
The \tool{ANTLR} grammar for the \cat language was extracted from the parser used by \tool{herd} tool%
\footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
written in OCaml.
%TODO: upload the grammar to github
%The input memory model 
\textbf{TODO: add reference to the grammar file in github repository OR include a short version as an appendix}

@@

@@

@@

<COmpiler contains several modules: (todo: remove todo about renaming interpreter to compiler: it's indeed interpreter>

@@

%todo: rename Interpreter -> Compiler.

>>>>

\begin{figure}[!h]%[!hb]
\centering
\begin{lstlisting}[language=antlr,
                  basicstyle=\ttfamily\tiny,
                  caption={The sketch of the input language grammar used by \oldporthos},%
                  label={fig:in_grammar_pts}]
<prog> : <init> <thrd>* <assert>
       ;
<thrd> : thread <tid> <inst>
       ;
<inst> : <atom>
       | <inst> ; <inst>
       | while <pred> <inst>
       | if <pred> { <inst> } <inst>
       ;
<atom> : <reg> <- <expr>
       | <reg> <:- <loc>
       | <loc> := <reg>
       | <reg> = <loc>.load(<atomic>)
       | <loc> = <reg>.store(<atomic>)
       | ('mfence' | 'sync' | 'lwsync' | 'isync')
       ;
<pred> :
       | true
       | false
       | <expr> (and | or) <expr>
       | <expr> ('==' | '!=' | '>' | '>=' | '<' | '>=') <expr>
       ;
<expr> : [0-9]
       | <reg>
       | <expr> ('*' | '+' | '-' | '/' | '%') <expr>
       ;
\end{lstlisting}
\end{figure}

@@

Following the litmus tests format, we distinguish three types of \xgraph{}: one for a process, one for a litmus-initialisation block and one for the assertion statement.
However, all three types of \xgraph{} are modelled by the same graph structure \texttt{XProcess} with certain restrictions complied by the corresponding type of X-interpreter that constructs the graph. %TODO: rename to XThread ?
This simplifies drastically the processing of different types of code blocks as they all are modelled by the same data structure.
The examples of restrictions on X-graphs are the following: the initialisation block can not have branchings, fence events; the process cannot have assertion events; the assertion block can not have shared-memory events or fence events.
We discuss the interpretation of different types of statements in more detail in Section~\ref{ch:impl:proc:x-compiler:compilation}.

@@

The input memory model language \cat{} is discussed in Section~\ref{ch:wmm:cat}.
The \tool{ANTLR} grammar for the \cat{} language was extracted from the parser used by \tool{herd} tool%
\footnote{\tool{herd} project repository: \url{https://github.com/herd/herdtools7}} %
written in OCaml.
%TODO: upload the grammar to github
%The input memory model 
\textbf{TODO: add reference to the grammar file in github repository OR include a short version as an appendix}

@@

- Y-Sytnax error

@@

\begin{figure}[th]
    \centering
  %draft=true
  \includegraphics[max width=\textwidth,keepaspectratio]{img/my/draw.io/X-compiler.png}
  \caption{Main components of the X-compilation processing unit}
  \label{fig:compiler}
\end{figure}

@@

The X-compiler is a complex processing unit, Figure~\ref{fig:compiler} illustrates the relationship the principal components of the X-compiler in the UML language.

The X-interpreter is a stateful processing unit that carries the \xgraph{}-builder and provides \textit{action} methods for changing its state and thus the state of the builder.

X compilation error

@@
